\LARGE \bf Geometric Adaptive Control of Attitude Dynamics on $\SO$\\ with State Inequality Constraints

### 2015-09-26 ACC16.2.tex
reduced down to 8 pages

\usepackage{ieee_packages}

@@

Attitude control is typically studied using a variety of attitude parameterizations, such as Euler angles or quaternions~\cite{shuster1993}.
All attitude parameterizations fail to represent the nonlinear configuration space both globally and uniquely~\cite{chaturvedi2011a}.
For example, minimal attitude representations, such as Euler angle sequences or modified Rodriguez parameters, suffer from singularities.
These attitude representations are not suitable for large angular slews.
%In order to avoid singularities, the designer must carefully switch the chosen Euler angle sequence based on the operating region.
%Another option is to artificially limit the operating region of the rigid body.
This ensures the system operates in a region free from singularities but limits the performance capabilities and ability to perform arbitrarily large angular maneuvers.
Quaternions do not have singularities but they double cover the special orthogonal group.
As a result, any physical attitude is represented by a pair of antipodal quaternions on the three-sphere.
%An immediate implication of this ambiguity is that closed-loop stability properties derived using quaternions may not hold for the physical rigid body evolving on the special orthogonal group.
During implementation, the designer must carefully resolve this non-unique representation in quaternion based attitude control systems to avoid undesirable unwinding behavior~\cite{bhat2000}, where the rigid body starts close to the desired attitude, yet the system unnecessarily rotates through a large angle even while the initial error is small.

% introduce concept of state inequality constraints and how it creates a nonconvex region 
% mention applicability of this for spacecraft maneuvers, spherical air bearing testing, hardware limitations
Many physical rigid body systems must perform large angular slews in the presence of state constraints.
For example, autonomous spacecraft or aerial systems are typically equipped with sensitive optical payloads, such as infrared or interferometric sensors.
These systems require retargeting while avoiding direct exposure to sunlight or other bright objects.
Determining a satisfactory attitude control maneuver in the presence of state constraints is a challenging task.
The removal of constrained regions from the rotational configuration space results in a \textit{nonconvex} region.
The attitude control problem in the absence of constraints has been extensively studied~\cite{bullo2004,MayTeePaCC11,LEEITAC15}.
However, the attitude control problem in the presence of constraints has received much less attention.

@@

% Previous potential function approach - developed using att parameterizations - singularities/ambiguities
Artificial potential functions are commonly used to handle kinematic constraints for a wide range of problems in robotics.
The goal is the design of an attractive function which drives the system towards the desired state.
Similarly, a repulsive function is constructed such that the system is directed away from the constraints. 
The superposition of the these functions allows one to apply standard feedback control schemes for stabilization and tracking.
Artificial potential functions have previously been applied to the attitude control problem in~\cite{lee2011b,mcinnes1994}.
These approaches were developed using attitude parameterizations, namely Euler angles and quaternions.
By characterizing the attitude both globally and uniquely \SO, we completely avoid issues associated with local parameterizations of the configuration manifold.

In short, the main contribution of this paper is developing an adaptive attitude control scheme in the presence of attitude inequality constraint on \SO.
We apply a potential function based approach developed directly on the nonlinear manifold \SO.
Our approach avoids the issues of attitude parameterizations, such as kinematic singularities and ambiguities, and is geometrically exact.
Our approach is also computationally efficient and able to handle uncertain disturbances. 
Instead of calculating a priori trajectories, as in the geometric and randomized approaches, our approach results in a closed-loop attitude control system.
This makes it ideal for on-board implentation on UAV or spacecraft systems.
In addition, unlike previous approaches our control system can handle an arbitrary number of constrained regions without modification. 
Finally, a rigous Lyapunov analysis guarantees the stability of the proposed control system, rather than probablistic convergence properties of the randomized approach.

We develop a new configuration error function on \SO with a logarithmic barrier function to avoid constrained regions.
We present a mathematically rigorous analysis to guarantee stability of the proposed control system.
In addition, we formulate an adaptive update law to enable attitude convergence in the presence of uncertain disturbances.
We demonstrate the effectiveness of this new approach via numerical simulation and experimental results on a hexrotor aerial vehicle.
%In this work, we develop a geometric adaptive controller on \(\SO\) to reorient a rigid body in the presence of state inequality constraints.

@@

\Cref{eqn:psi} is composed of an attractive term, \( A (R) \), toward the desired attitude and a repulsive term, \( B(R) \), away from the undesired direction \( R^T v \).
The attractive error function has been used for attitude control on \SO.
The potential well of \( A(R)\) is illustrated in~\cref{fig:attract_error}.
The desired attitude lies at the minimum of \( A(R) \).

@@

While~\cref{eqn:B} represents a single inequality constraint given as~\cref{eqn:constraint}, it is readily generalized to multiple constraints of an arbitrary form. 
For example, the configuration error function for multiple constraints can be formulated as $\Psi=A[1+\sum_i C_i]$, where $C_i$ has the form of $C_i=B-1$ for the $i$-th constraint. 
In this manner, one may enforce multiple state inequality constraints on the control system.

\begin{prop}[Error Dynamics]\label{prop:error_dyn}
	The attitude error dynamics for \( \Psi, e_R, e_\Omega \) satisfy 
	\begin{gather}
		\diff{}{t} \parenth{\Psi} = e_R \cdot e_\Omega \, , \label{eqn:psi_dot}\\
		\diff{}{t} \parenth{e_R} = \dot{e}_{R_A} B + e_{R_A} \dot{B} + \dot{A}e_{R_B} + A \dot{e}_{R_B} \, , \label{eqn:eR_dot}\\
		\diff{}{t} \parenth{e_{R_A}} = E(R, R_d) e_\Omega \, , \label{eqn:eRA_dot} \\
		\diff{}{t} \parenth{e_{R_B}} = F(R) e_\Omega \, , \label{eqn:eRB_dot} \\
		\diff{}{t} \parenth{A} = e_{R_A} \cdot e_\Omega \, , \label{eqn:A_dot} \\
		\diff{}{t} \parenth{B} = e_{R_B} \cdot e_\Omega \, , \label{eqn:B_dot} \\
		\diff{}{t} \parenth{e_\Omega} = J^{-1} \parenth{-\Omega \times J \Omega + u + W(R, \Omega) \Delta} \, , \label{eqn:eW_dot}
	\end{gather}
	where the matrices \(E(R,R_d), F(R) \in \R^{3\times3} \) are given by
	\begin{gather}
		E(R,R_d) = \frac{1}{2} \parenth{\tr{R^T R_d G}I - R^T R_d G} \, , \label{eqn:E} \\
		F(R) = \frac{1}{\alpha \parenth{r^T R^T v - \cos \theta}} \left[\parenth{v^T R r} I - R^T v r^T + \right. \nonumber \\
		\left. \frac{R^T \hat{v} R r v^T R \hat{r}}{\parenth{r^T R^T v - \cos \theta}}\right] \, . \label{eqn:F}
	\end{gather}
\end{prop}
\begin{proof}
See~\Cref{proof:error_dyn}.
\end{proof}

@@

%This results in \( A \) having a value of \( g_1 + g_2, g_2 + g_3 \) or \( g_3 + g_1 \).
%At any of these undesired equilibria we have \( V = k \). 
%We define an augmented Lyapunov function as
%\begin{align*}
%	W = k - V = \frac{1}{2} e_\Omega \cdot J e_\Omega + k_R \parenth{\frac{k}{k_R}+ \Psi(R,R_d)} \, .
%\end{align*}
%We can choose an attitude \( R \) arbitrarily close to the equilibrium point \( \exp{ \parenth{\pi \hat{e}_i}} R_d \) such that the \( k_R \parenth{\frac{k}{k_R}+\Psi }> 0 \).
%This is due to the fact that \( \Psi \) is positive definite about \( R = R_d \) as shown in~\cref{item:prop_psi_psd}.
%If \( e_\Omega \) is sufficiently small this results in \( W > 0 \) and \( \dot{W} = -\dot{V} > 0\) from~\cref{eqn:vdot_nodist}.
%According to Chateav's theorem, the undesired equilibria are unstable~\cite{khalil1996}.
\subsection{Adaptive Control}
We extend the results of the previous section with the addition of a fixed but unknown disturbance \( \Delta \).
This scenario is typical of many mechanical systems and represents unmodeled dynamics or external moments acting on the system.
For example, Earth orbiting spacecraft typically experience both a torque due to a gravitational gradient as well as external torques to due solar radiation pressure.
Aerial vehicles will similarly experience external torques due to air currents or turbulence.
An adaptive control system is introduced to asymptotically stabilize the system to a desired attitude while ensuring that state constraints are satisfied.

@@

Compared to the previous work on constrained attitude control, we present a geometrically exact control system without parameterizations.
In addition, we incorporate state inequality constraints for the first time on \( \SO \).
The presented control system is computed in real-time and offers significant computational advantages over previous iterative methods. 
In addition, the riguous mathematical proof guarantees stability.
\section{Numerical Examples}
We demonstrate the performance of the proposed control system via numerical simulation.
The inertia tensor of a rigid body is given as
\begin{gather*}
	J = \begin{bmatrix}
	\num{5.57e-3} & \num{6.17e-5} & \num{-2.50e-5} \\
	\num{6.17e-5} & \num{5.57e-3} & \num{1.00e-5} \\
	\num{-2.50e-5} & \num{1.00e-5} & \num{1.05e-2}
	\end{bmatrix} \si{\kilo\gram\meter\squared} \, .
\end{gather*} 
The control system parameters are chosen as
\begin{gather*}
	G = \text{diag} \parenth{0.9,1.1,1.0}\, , \quad k_R = 0.4 \, , \quad	k_\Omega = 0.296 \, ,\\
	c = 1.0 \, , \quad k_\Delta = 0.5 \, , \quad \alpha = 15 \, .
\end{gather*}
A body fixed sensor is defined as \(r = \begin{bmatrix} 1 & 0 & 0 \end{bmatrix}\), while multiple inequality constraints are defined in~\Cref{tab:constraints}.
The simulation parameters are chosen to be similar to those found in~\cite{lee2011b}, howver we increase the size of the constraint regions to create a more challenging scenario for the control system.
The initial state is defined as \(R_0 =  \exp(\ang{225} \times \frac{\pi}{180} \hat{e}_3), \Omega_0 = 0\). 
The desired state is \( (R_d = I,\Omega_d = 0)\).
\begin{table}[htbp]
\caption{Constraint Parameters~\label{tab:constraints}}
\begin{center}\begin{tabular}{lc}
Constraint Vector (\( v \)) & Angle (\( \theta \)) \\ \hline \hline 
\(\begin{bmatrix} 0.174 &-0.934 & -0.034 \end{bmatrix}\) & \ang{40} \\ \hline 
\(\begin{bmatrix} 0 & 0.7071 & 0.7071 \end{bmatrix}\) & \ang{40} \\ \hline 
\(\begin{bmatrix} -0.853 & 0.436 & -0.286 \end{bmatrix}\) & \ang{40} \\ \hline 
\(\begin{bmatrix} -0.122 -0.140 -0.983 \end{bmatrix}\) & \ang{20}\end{tabular} 
\end{center}
\end{table}
We show simulation results for the system stabilizing about the desired attitude with and without the adaptive update law from~\Cref{prop:adaptive_control}.
We assume a fixed disturbance of \(\Delta = \begin{bmatrix} 0.2 & 0.2 & 0.2 \end{bmatrix}^T \), with the function \( W(R,\Omega) = I \).
This form is equivalent to an integral control term which penalizes deviations from the desired configuration.
The first term of~\cref{eqn:delta_dot} has the effect of increasing the proportional gain of the control system, since the time derivative of the attitude error vector, \( \dot{e_{R}} \), is linear with respect to the angular velocity error vector \( e_\Omega\).
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{eR_con} 
		\caption{Attitude error vector \(e_R\) } \label{fig:eR_con} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{Psi_con} 
		\caption{Configuration error \( \Psi \)} \label{fig:Psi_con} 
	\end{subfigure}
	\caption{Attitude stabilization without adaptive update law}
	\label{fig:con} 
\end{figure}

\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{eR_adapt} 
		\caption{Attitude error vector \(e_R\) } \label{fig:eR_adapt} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{Psi_adapt} 
		\caption{Configuration error \( \Psi \)} \label{fig:Psi_adapt} 
	\end{subfigure}

@@

Simulation results without the adaptive update law are shown in~\cref{fig:con}.
Without the update law, the system does not achieve zero steady state error. 
\Cref{fig:Psi_con} shows that the configuration error function does not converge to zero and there exist steady state errors.
\Cref{fig:adapt} shows the results with the addition of the adaptive update law.
The addition of the adaptive update law allows the system to converge to the desired attitude in the presence of constraints.
The path of the body fixed sensor in the inertial frame, namely \( R r \), is illustrated in~\cref{fig:cad_adapt}.
The initial attitude is represented with the green circle while the final attitude is marked with a green \(\times\).
The inequality constraints from~\Cref{tab:constraints} are depicted as red cones, where the cone half angle is \( \theta \).
The control system is able to asymptotically converge to zero attitude error.
In addition, the estimate of the disturbance converges to the the true value.

Both control system are able to automatically avoid the constrained regions. 
In addition, these results show that it is straightforward to incorporate an arbitrary amount of large constraints.
In spite of this challenging configuration space the proposed control system offers a simple method of avoiding constrained regions.
These closed-loop feedback results are computed in real time and offer a significant advantage over typical open-loop planning methods.
These results show that the proposed geometric adaptive approach is critical to attitude stabilization in the presence of state constraints and disturbances.
\section{Experiment on Hexrotor UAV}
A hexrotor unmanned aerial vehicle (UAV) has been developed at the Flight Dynamics and Controls Laboratory(FDCL) at the George Washington University~\cite{kaufman2014}.
The UAV is composed of three pairs of counter-rotating propellers. 
Typical quadrotor UAVs are composed of four or more co-planar propellers.
As a result, these systems are underactuated and unable to impart a force along every degree of freedom.
For example, quadrotor UAVs are unable to translate laterally without first conducting a rotation.
Conversely, the propeller pairs of the hexrotor are angled relative to one another to allow for a fully actuated rigid body.
This allows the hexrotor to impart a force in any direction and a moment about any axis.

The hexrotor UAV, shown in~\cref{fig:hexrotor}, is composed of the following hardware:
\begin{itemize}
	\item Onboard ODROID XU3 computer module.
	\item VectorNav VN100 IMU operating at over \SI{100}{\hertz} via TTL serial 
	\item Brushless DC Motors powered by BL-Ctrl-2.0 speed controllers via I2C.
	\item Position and attitude over WiFi (TCP) communication from Vicon motion capture system via host computer.
	\item Commands sent over WiFi through the terminal (SSH) to the code running onboard the hexrotor. 
\end{itemize}
In order to constrain the motion and test only the attitude dynamics we attach the hexrotor to a spherical joint.
The center of rotation is below the center of gravity of the hexrotor.
As a result, there is a destabilizing gravitational moment and the resulting attitude dynamics are similar to an inverted pendulum model.
We augment the control input in~\cref{eqn:adaptive_control} with an additional term to negate the effect of the gravitational moment.

>>>>

\usepackage{ieee_packages}
\newcommand{\squeezeup}{\vspace{-2.5mm}}

@@

Attitude control is typically studied using a variety of attitude parameterizations, such as Euler angles or quaternions~\cite{shuster1993}.
All attitude parameterizations fail to represent the nonlinear configuration space both globally and uniquely~\cite{chaturvedi2011a}.
For example, minimal attitude representations, such as Euler angle sequences or modified Rodriguez parameters, suffer from singularities.
These attitude representations are not suitable for large angular slews.
%In order to avoid singularities, the designer must carefully switch the chosen Euler angle sequence based on the operating region.
%Another option is to artificially limit the operating region of the rigid body.
%This ensures the system operates in a region free from singularities but limits the performance capabilities and ability to perform arbitrarily large angular maneuvers.
Quaternions do not have singularities but they double cover the special orthogonal group.
As a result, any physical attitude is represented by a pair of antipodal quaternions on the three-sphere.
%An immediate implication of this ambiguity is that closed-loop stability properties derived using quaternions may not hold for the physical rigid body evolving on the special orthogonal group.
During implementation, the designer must carefully resolve this non-unique representation in quaternion based attitude control systems to avoid undesirable unwinding behavior~\cite{bhat2000}.
%, where the rigid body starts close to the desired attitude, yet the system unnecessarily rotates through a large angle even while the initial error is small.

% introduce concept of state inequality constraints and how it creates a nonconvex region 
% mention applicability of this for spacecraft maneuvers, spherical air bearing testing, hardware limitations
Many physical rigid body systems must perform large angular slews in the presence of state constraints.
For example, autonomous spacecraft or aerial systems are typically equipped with sensitive optical payloads, such as infrared or interferometric sensors.
These systems require retargeting while avoiding direct exposure to sunlight or other bright objects.
%Determining a satisfactory attitude control maneuver in the presence of state constraints is a challenging task.
The removal of constrained regions from the rotational configuration space results in a \textit{nonconvex} region.
The attitude control problem in the absence of constraints has been extensively studied~\cite{bullo2004,MayTeePaCC11,LEEITAC15}.
However, the attitude control problem in the presence of constraints has received much less attention.

@@

% Previous potential function approach - developed using att parameterizations - singularities/ambiguities
Artificial potential functions are commonly used to handle kinematic constraints for a wide range of problems in robotics.
The goal is the design of attractive and repulsive terms which drive the system toward or away from a certain state, respectively.
%The goal is the design of an attractive function which drives the system towards the desired state.
%Similarly, a repulsive function is constructed such that the system is directed away from the constraints. 
The superposition of the these functions allows one to apply standard feedback control schemes for stabilization and tracking.
Artificial potential functions have previously been applied to the attitude control problem in~\cite{lee2011b,mcinnes1994}.
These approaches were developed using attitude parameterizations, namely Euler angles and quaternions.
By characterizing the attitude both globally and uniquely \(\SO\), we completely avoid issues associated with local parameterizations of the configuration manifold.

In short, the main contribution of this paper is developing an adaptive attitude control scheme in the presence of attitude inequality constraint on \(\SO\).
We apply a potential function based approach developed directly on the nonlinear manifold \(\SO\).
Our approach avoids the issues of attitude parameterizations, such as kinematic singularities and ambiguities, and is geometrically exact.
Our approach is also computationally efficient and able to handle uncertain disturbances. 
Instead of calculating a priori trajectories, as in the geometric and randomized approaches, our approach results in a closed-loop attitude control system.
This makes it ideal for on-board implementation on UAV or spacecraft systems.
In addition, unlike previous approaches our control system can handle an arbitrary number of constrained regions without modification. 
Finally, a rigorous Lyapunov analysis guarantees the stability of the proposed control system.
%rather than probablistic convergence properties of the randomized approach.

We develop a new configuration error function on \(\SO\) with a logarithmic barrier function to avoid constrained regions.
We present a mathematically rigorous analysis to guarantee stability of the proposed control system.
In addition, we formulate an adaptive update law to enable attitude convergence in the presence of uncertain disturbances.
We demonstrate the effectiveness of this new approach via numerical simulation and experimental results on a hexrotor aerial vehicle.
%In this work, we develop a geometric adaptive controller on \(\SO\) to reorient a rigid body in the presence of state inequality constraints.

@@

\Cref{eqn:psi} is composed of an attractive term, \( A (R) \), toward the desired attitude and a repulsive term, \( B(R) \), away from the undesired direction \( R^T v \).
The attractive error function has been used for attitude control on \(\SO\).
The potential well of \( A(R)\) is illustrated in~\cref{fig:attract_error}.
The desired attitude lies at the minimum of \( A(R) \).

@@

While~\cref{eqn:B} represents a single inequality constraint given as~\cref{eqn:constraint}, it is readily generalized to multiple constraints of an arbitrary form. 
For example, the configuration error function for multiple constraints can be formulated as $\Psi=A[1+\sum_i C_i]$, where $C_i$ has the form of $C_i=B-1$ for the $i$-th constraint. 
In this manner, one may enforce multiple state inequality constraints on the control system.
\begin{prop}[Error Dynamics]\label{prop:error_dyn}
	The attitude error dynamics for \( \Psi, e_R, e_\Omega \) satisfy 
	\begin{gather}
		\diff{}{t} \parenth{\Psi} = e_R \cdot e_\Omega \, , \label{eqn:psi_dot}\\
		\diff{}{t} \parenth{e_R} = \dot{e}_{R_A} B + e_{R_A} \dot{B} + \dot{A}e_{R_B} + A \dot{e}_{R_B} \, , \label{eqn:eR_dot}\\
		\diff{}{t} \parenth{e_{R_A}} = E(R, R_d) e_\Omega \, , \label{eqn:eRA_dot} \\
		\diff{}{t} \parenth{e_{R_B}} = F(R) e_\Omega \, , \label{eqn:eRB_dot} \\
		\diff{}{t} \parenth{A} = e_{R_A} \cdot e_\Omega \, , \label{eqn:A_dot} \\
		\diff{}{t} \parenth{B} = e_{R_B} \cdot e_\Omega \, , \label{eqn:B_dot} \\
		\diff{}{t} \parenth{e_\Omega} = J^{-1} \parenth{-\Omega \times J \Omega + u + W(R, \Omega) \Delta} \, , \label{eqn:eW_dot}
	\end{gather}
	where the matrices \(E(R,R_d), F(R) \in \R^{3\times3} \) are given by
	\begin{gather}
		E(R,R_d) = \frac{1}{2} \parenth{\tr{R^T R_d G}I - R^T R_d G} \, , \label{eqn:E} \\
		F(R) = \frac{1}{\alpha \parenth{r^T R^T v - \cos \theta}} \left[\parenth{v^T R r} I - R^T v r^T + \right. \nonumber \\
		\left. \frac{R^T \hat{v} R r v^T R \hat{r}}{\parenth{r^T R^T v - \cos \theta}}\right] \, . \label{eqn:F}
	\end{gather}
\end{prop}
\begin{proof}
See~\Cref{proof:error_dyn}.
\end{proof}

@@

%This results in \( A \) having a value of \( g_1 + g_2, g_2 + g_3 \) or \( g_3 + g_1 \).
%At any of these undesired equilibria we have \( V = k \). 
%We define an augmented Lyapunov function as
%\begin{align*}
%	W = k - V = \frac{1}{2} e_\Omega \cdot J e_\Omega + k_R \parenth{\frac{k}{k_R}+ \Psi(R,R_d)} \, .
%\end{align*}
%We can choose an attitude \( R \) arbitrarily close to the equilibrium point \( \exp{ \parenth{\pi \hat{e}_i}} R_d \) such that the \( k_R \parenth{\frac{k}{k_R}+\Psi }> 0 \).
%This is due to the fact that \( \Psi \) is positive definite about \( R = R_d \) as shown in~\cref{item:prop_psi_psd}.
%If \( e_\Omega \) is sufficiently small this results in \( W > 0 \) and \( \dot{W} = -\dot{V} > 0\) from~\cref{eqn:vdot_nodist}.
%According to Chateav's theorem, the undesired equilibria are unstable~\cite{khalil1996}.
\subsection{Adaptive Control}
We extend the results of the previous section with the addition of a fixed but unknown disturbance \( \Delta \).
This scenario is typical of many mechanical systems and represents unmodelled dynamics or external moments acting on the system.
For example, Earth orbiting spacecraft typically experience both a torque due to a gravitational gradient.
% as well as external torques to due solar radiation pressure.
Aerial vehicles will similarly experience external torques due to air currents or turbulence.
An adaptive control system is introduced to asymptotically stabilize the system to a desired attitude while ensuring that state constraints are satisfied.

@@

Compared to the previous work on constrained attitude control, we present a geometrically exact control system without parameterizations.
In addition, we incorporate state inequality constraints for the first time on \( \SO \).
The presented control system is computed in real-time and offers significant computational advantages over previous iterative methods. 
In addition, the riguous mathematical proof guarantees stability.
\section{Numerical Examples}
We demonstrate the performance of the proposed control system via numerical simulation.
The inertia tensor of a rigid body is given as
\begin{gather*}
	J = \begin{bmatrix}
	\num{5.57e-3} & \num{6.17e-5} & \num{-2.50e-5} \\
	\num{6.17e-5} & \num{5.57e-3} & \num{1.00e-5} \\
	\num{-2.50e-5} & \num{1.00e-5} & \num{1.05e-2}
	\end{bmatrix} \si{\kilo\gram\meter\squared} \, .
\end{gather*} 
The control system parameters are chosen as
\begin{gather*}
	G = \text{diag} \parenth{0.9,1.1,1.0}\, , \quad k_R = 0.4 \, , \quad	k_\Omega = 0.296 \, ,\\
	c = 1.0 \, , \quad k_\Delta = 0.5 \, , \quad \alpha = 15 \, .
\end{gather*}
A body fixed sensor is defined as \(r = \begin{bmatrix} 1 & 0 & 0 \end{bmatrix}\), while multiple inequality constraints are defined in~\Cref{tab:constraints}.
The simulation parameters are chosen to be similar to those found in~\cite{lee2011b}, howver we increase the size of the constraint regions to create a more challenging scenario for the control system.
The initial state is defined as \(R_0 =  \exp(\ang{225} \times \frac{\pi}{180} \hat{e}_3), \Omega_0 = 0\). 
The desired state is \( (R_d = I,\Omega_d = 0)\).
\begin{table}[htbp]
\caption{Constraint Parameters~\label{tab:constraints}}
\begin{center}\begin{tabular}{lc}
Constraint Vector (\( v \)) & Angle (\( \theta \)) \\ \hline \hline 
\(\begin{bmatrix} 0.174 &-0.934 & -0.034 \end{bmatrix}\) & \ang{40} \\ \hline 
\(\begin{bmatrix} 0 & 0.7071 & 0.7071 \end{bmatrix}\) & \ang{40} \\ \hline 
\(\begin{bmatrix} -0.853 & 0.436 & -0.286 \end{bmatrix}\) & \ang{40} \\ \hline 
\(\begin{bmatrix} -0.122 -0.140 -0.983 \end{bmatrix}\) & \ang{20}\end{tabular} 
\end{center}
\end{table}
We show simulation results for the system stabilizing about the desired attitude with and without the adaptive update law from~\Cref{prop:adaptive_control}.
We assume a fixed disturbance of \(\Delta = \begin{bmatrix} 0.2 & 0.2 & 0.2 \end{bmatrix}^T \), with the function \( W(R,\Omega) = I \).
This form is equivalent to an integral control term which penalizes deviations from the desired configuration.
The first term of~\cref{eqn:delta_dot} has the effect of increasing the proportional gain of the control system, since the time derivative of the attitude error vector, \( \dot{e_{R}} \), is linear with respect to the angular velocity error vector \( e_\Omega\).
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{eR_con} 
		\caption{Attitude error vector \(e_R\) } \label{fig:eR_con} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{Psi_con} 
		\caption{Configuration error \( \Psi \)} \label{fig:Psi_con} 
	\end{subfigure}
	\caption{Attitude stabilization without adaptive update law}
	\label{fig:con} 
\end{figure}
\begin{figure} 
	\centering 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{eR_adapt} 
		\caption{Attitude error vector \(e_R\) } \label{fig:eR_adapt} 
	\end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
	\begin{subfigure}[htbp]{0.5\columnwidth} 
		\includegraphics[width=\columnwidth]{Psi_adapt} 
		\caption{Configuration error \( \Psi \)} \label{fig:Psi_adapt} 
	\end{subfigure}

@@

Simulation results without the adaptive update law are shown in~\cref{fig:con}.
Without the update law, the system does not achieve zero steady state error. 
\Cref{fig:Psi_con} shows that the configuration error function does not converge to zero and there exist steady state errors.
\Cref{fig:adapt} shows the results with the addition of the adaptive update law.
The addition of the adaptive update law allows the system to converge to the desired attitude in the presence of constraints.
The path of the body fixed sensor in the inertial frame, namely \( R r \), is illustrated in~\cref{fig:cad_adapt}.
The initial attitude is represented with the green circle while the final attitude is marked with a green \(\times\).
The inequality constraints from~\Cref{tab:constraints} are depicted as red cones, where the cone half angle is \( \theta \).
The control system is able to asymptotically converge to zero attitude error 
In addition, the estimate of the disturbance converges to the the true value.

Both control system are able to automatically avoid the constrained regions. 
In addition, these results show that it is straightforward to incorporate an arbitrary amount of large constraints.
In spite of this challenging configuration space the proposed control system offers a simple method of avoiding constrained regions.
These closed-loop feedback results are computed in real time and offer a significant advantage over typical open-loop planning methods.
These results show that the proposed geometric adaptive approach is critical to attitude stabilization in the presence of state constraints and disturbances.
\section{Experiment on Hexrotor UAV}
A hexrotor unmanned aerial vehicle (UAV) has been developed at the Flight Dynamics and Controls Laboratory(FDCL) at the George Washington University~\cite{kaufman2014}.
The UAV is composed of three pairs of counter-rotating propellers. 
%Typical quadrotor UAVs are composed of four or more co-planar propellers.
%As a result, these systems are underactuated and unable to impart a force along every degree of freedom.
%For example, quadrotor UAVs are unable to translate laterally without first conducting a rotation.
%Conversely, the propeller pairs of the hexrotor are angled relative to one another to allow for a fully actuated rigid body.
The propeller pairs of the hexrotor are angled relative to one another to allow for a fully actuated rigid body.
%This allows the hexrotor to impart a force in any direction and a moment about any axis.

The hexrotor UAV, shown in~\cref{fig:hexrotor}, is composed of the following hardware:
\begin{itemize}
	\item Onboard ODROID XU3 computer module.
	\item VectorNav VN100 IMU operating via TTL serial %at over \SI{100}{\hertz} 
	\item BLDC motors with BL-Ctrl-2.0 ESC via I2C.
	\item Position and attitude over WiFi (TCP) communication from Vicon motion capture system.
	\item Commands sent over WiFi to  onboard controller. 
\end{itemize}
In order to constrain the motion and test only the attitude dynamics we attach the hexrotor to a spherical joint.
The center of rotation is below the center of gravity of the hexrotor.
As a result, there is a destabilizing gravitational moment and the resulting attitude dynamics are similar to an inverted pendulum model.
We augment the control input in~\cref{eqn:adaptive_control} with an additional term to negate the effect of the gravitational moment.
